# Usage 

## Running the Model - Basic

To run the calorie model you must have [R](https://cran.rstudio.com/) and [RStudio](https://rstudio.com/products/rstudio/download/#download) installed (if in DHSC you can do this via a standard [software request](https://in.dh.gov.uk/how-to/seek-it-support/it-help-and-support/) and also must fix projects by running the solution in [section C4](https://datas-dhsc.github.io/coding_principles_book/r-at-dhsc.html#r_projects))

1. Copy the folder `\\ims.gov.uk\DHSC\London\SKH\NW012\EOR3\COP\Calorie Model\Calorie Model `r version_number`` to your local machine and rename it as a folder called `[Policy]-[yyyymmdd]-[analyst name]`. For example, if Edward Beake was running a model simulation for the advertising policy on 21/10/2019 the copy of the folder `Calorie Model `r version_number`` would be renamed `Advertising-20191021-EBeake`. (This will take a few minutes).

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)

src_files <- list.files("../../Calorie Model 3/", recursive = T, full.names = T,
                       pattern = "\\.R$")


find_libraries <- function(file){
 

  read_lines(file) %>%
    str_subset("(library|require)") %>%
    str_remove("(library|require)") %>%
    str_extract("[a-zA-Z0-9]{3,}")
  
  
}
libraries <- map(src_files,find_libraries) %>%
  unlist() %>%
  unique() %>%
  str_subset("\\w") %>%
  str_replace_all("(^|$)", "'") %>%
  str_flatten(collapse = " , ")

install_string <- str_glue("install.packages(c(", libraries,"))")
```

2. Install the required packages by running this code in the R console: **`r install_string`**.

3. Open the R project file `Calorie Model `r version_number`.Rproj` in the renamed folder.

4. From the `/src/Model code` sub folder open the main code file `Calorie Model `r version_number`.R`, if this code file is not open.

5. Run the model with `Ctrl+1` then `Ctrl+A+ENTER`

6. Follow the on screen instructions in the pop up boxes. These pop up boxes allow the user to change all standard model parameters, before starting a model run you should have answers to all these questions:

    * do the model cohorts include adults,

    * what is the evaluation period of the policy,

    * what is the lifetime of the policy (this is the length of time the policy is active and effects new model cohorts, once affected by the policy a cohort is permantly affected - for further information see [Policy Permanence]),
    
    * is the NHS oppertunity cost multiplier used (a 4 x multiplier applied to NHS costs),

    * what are the calorie reduction for the 8 standard groups:

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.width=5}
 kable(data.frame(Sex = c('Male',
                           'Male',
                           'Male',
                           'Male',
                           'Female',
                           'Female',
                          'Female',
                          'Female'),
                  Age.Catagory = c('4 - 10',
                                   '11 - 18',
                                   '19 - 64',
                                   '65+',
                                   '4 - 10',
                                   '11 - 18',
                                   '19 - 64',
                                   '65+'))) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 12,full_width = FALSE) 
```

For more advanced functionality, such as using different calorie reduction age groups, changes will need to be 'hard coded' into the model and the model developer consulted.

A full model run will take approximatly 20 minutes, the exact time the model takes will depend on the parameters specified in step 6 (NOTE: Some warning messages - these appear in red on the R console - are expected and should not concern the user).

After the model is finished:

* The model will be saved, using the name specified in the last pop up box of step 6, as an `.rds` file in the new sub-folder `model`. This file can be loaded into R and contains the complete details of your model.

* The model summary report will be saved in the new sub-folder `report`.

* Archive the entire folder `[Policy]-[yyyymmdd]-[analyst name]` by moving it to the archive folder `\\ims.gov.uk\DHSC\London\SKH\NW012\EOR3\COP\Calorie Model\Model Run Archive\[Policy]`, this is done to maintain a full audit trail of the model run.

## Running the Model - Advanced

Advanced runs of the model can be done by hard coding into the file `Calorie Model `r version_number`.R`.

## Developing the Model

The majority of model developments, updating states and transition probabilities to add new diseases or change the obesity related prevalence increase, are made in the [model input files][Model Parameters and Data -`/heemod data`].
